### 2021-07-15 src/figures/results/tables/student_infections_table.tex

\begin{tabular}{lr}
\toprule
{} &  predicted total infections among 5-14 year olds from Easter until 2021-05-31 \\
scenario                               &                                                                               \\
\midrule
 educ open after easter  without tests &                                             696963 \\
 educ open after easter  with tests    &                                             545057 \\
 close educ after easter               &                                             422360 \\
 baseline                              &                                             432320 \\
\bottomrule
\end{tabular}

>>>>

\begin{tabular}{lr}
\toprule
{} &  predicted total infections among 5-14 year olds from Easter until 2021-05-31 \\
scenario                               &                                                                               \\
\midrule
 educ open after easter  without tests &                                             951185 \\
 educ open after easter  with tests    &                                             722502 \\
 close educ after easter               &                                             559526 \\
 baseline                              &                                             582466 \\
\bottomrule
\end{tabular}

### 2021-07-15 src/paper/abstract.tex

\noindent
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially,
these focused on non-pharmaceutical interventions; more recently, vaccinations and
large-scale rapid testing have started to play a major role. The objective of this study
is to explain the quantitative effects of these policies on determining the course of
the pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation model,
which is estimated using data for the second and the third wave of the CoViD-19 pandemic
in Germany. The paper finds that during a period where vaccination rates rose from 5\%
to 40\%, rapid testing had the largest effect on reducing infection numbers. Frequent
large-scale rapid testing should remain part of strategies to contain CoViD-19; it can
substitute for many non-pharmaceutical interventions that come at a much larger cost to
individuals, society, and the economy.

>>>>

\noindent
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially, these
focused on non-pharmaceutical interventions; more recently, vaccinations and large-scale
rapid testing have started to play a major role. The objective of this study is to
explain the quantitative effects of these policies on determining the course of the
pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation model,
which is estimated using data for the second and the third wave of the CoViD-19 pandemic
in Germany. The paper finds that during a period where vaccination rates rose from 5\% to
40\%, seasonality and rapid testing had the largest effect on reducing infection numbers.
Frequent large-scale rapid testing should remain part of strategies to contain CoViD-19;
it can substitute for many non-pharmaceutical interventions that come at a much larger
cost to individuals, society, and the economy.

### 2021-07-15 src/paper/data/number_of_contacts.tex

    \caption{Number of Contacts of the Different Contact Types}
    \label{fig:n_contacts}
    \floatfoot{\noindent
        \textit{Note:} This figure shows the pre-pandemic number of contacts individuals
        report of different contact types ($\eta_c$).
        % non recurrent contacts
        In the model it is sampled every day which of the numbers of non recurrent
        contacts a person is planned to have. Note that the contact diaries include such
        high values that super spreading events are well possible in our model through
        non recurrent models.
        % recurrent contacts
        For recurrent contacts individuals are put into groups that meet either every day
        or on a particular week day every day. For work contacts, meetings can only take
        place on work days.
        % reductions that can happen
        The pre-pandemic number of contacts with transmission potential is reduced by
        policies ($\rho$), seasonality ($\kappa$) and individual responses ($\tau$) to
        events such as receiving a positive rapid test to the number of actual contacts
        with transmission potential.
        % other
        The upper row shows the distribution of the number of other contacts individuals
        report ($\eta_{other}$). Other contacts include all contacts that are not
        household members, school contacts or work contacts, for example leisure
        contacts. We assume that individuals in households with children or teachers or
        retired individuals have additional non recurrent other contacts during school
        vacations to cover things like family visits or travel during vacations.
        % work contacts
        The lower row shows the distribution of the different types of work contacts
        ($\eta_{work}$). Work contacts only take place between working individuals.}
\end{figure}

>>>>

    \caption{Number of Contacts of the Different Contact Types}
    \label{fig:n_contacts}
    \floatfoot{\noindent
        \textit{Note:} This figure shows the pre-pandemic number of contacts individuals
        report of different contact types ($\eta_c$).
        % non recurrent contacts
        In the model it is sampled every day which of the numbers of non recurrent
        contacts a person is planned to have. Note that the contact diaries include such
        high values that super spreading events are well possible in our model through
        non recurrent models.
        % recurrent contacts
        For recurrent contacts individuals are put into groups that meet either every day
        or on a particular week day every day.
        % For work contacts, meetings can only take place on work days.
        % % reductions that can happen
        % The pre-pandemic number of contacts with transmission potential is reduced by
        % policies ($\rho$), seasonality ($\kappa$) and individual responses ($\tau$) to
        % events such as receiving a positive rapid test to the number of actual contacts
        % with transmission potential.
        % other
        The upper row shows the distribution of the number of other contacts individuals
        report ($\eta_{other}$). Other contacts include all contacts that are not
        household members, school contacts or work contacts, for example leisure
        contacts. We assume that individuals in households with children or teachers or
        retired individuals have additional non recurrent other contacts during school
        vacations to cover things like family visits or travel during vacations.
        % work contacts
        The lower row shows the distribution of the different types of work contacts
        ($\eta_{work}$). Work contacts only take place between working individuals.}
\end{figure}

### 2021-07-15 src/paper/data/testing_and_behavioral_response.tex

 \comment[id=K]{@Janos: Read this}

After showing how we calibrate rapid test demand and the share of detected cases, this
section describes the remaining parameters for our testing model. Refer to
Section~\ref{sub:testing} for a description of the full testing model.

% share of tests for symptomatics
From the share of detected cases and the number of infections we arrive at the number of
positive PCR tests in our model. A share of these positive tests is given to symptomatic
individuals ($\chi_{symptom,\:t}$). This share is calibrated from German data on case
characteristics \citep{RKICaseCharacteristics} and shown in
Figure~\ref{fig:share_pcr_to_symptomatic}. We keep $\chi_{symptom,\:t}$ constant after
Christmas because the RKI data does not include if a PCR test was done to confirm a
positive rapid test and this share is used for PCR test demand without prior rapid test
indication

@@

% share with positive rapid test requesting pcr
To model the demand for PCR tests through rapid tests, we only need the share of
individuals that seek a PCR test to confirm a positive rapid test result
($\chi_{confirmation}$). We calibrate this from \citet{Betsch2021} who asked this
hypothetical question in March of 2021. There 82\% of Germans reported that they would
follow up on a positive rapid test with a PCR test.

>>>>

This section describes the remaining parameters for our testing model. Refer to
Section~\ref{sub:testing} for a description of the full testing model.

% share of tests for symptomatics
From the share of detected cases and the number of infections we arrive at the number of
positive PCR tests in our model. A share of these positive tests goes to symptomatic
individuals ($\chi_{symptom,\:t}$). This share is calibrated from German data on case
characteristics \citep{RKICaseCharacteristics} and shown in
Figure~\ref{fig:share_pcr_to_symptomatic}. We keep $\chi_{symptom,\:t}$ constant after
Christmas because the RKI data does not include if a PCR test was done to confirm a
positive rapid test and this share is used for PCR test demand without prior rapid test
indication.

@@

% share with positive rapid test requesting pcr
To model the demand for PCR tests through rapid tests, we only need the share of
individuals that seek a PCR test to confirm a positive rapid test result
($\chi_{confirmation}$). We calibrate this from \citet{Betsch2021} who asked this as a
hypothetical question in March of 2021. There 82\% of Germans reported that they would
follow up on a positive rapid test with a PCR test.

### 2021-07-15 src/paper/literature_review.tex

These shortcomings have motivated variations of agent-based simulation models where
moving particles have been replaced by contact networks for households, work and random
contacts.

The OpenABM-Covid-19 model by \citet{Hinch2020} is the closest in spirit to
ours\comment[id=HM]{I'd rather be close to Aleta et al. ;-)}. They model detailed
contact networks for workplaces, schools and households and can evaluate the effect of
several NPIs. The main focus of their application are contact tracing policies.

@@

Moreover, there are several forthcoming applications\comment[id=HM]{What is that
supposed to mean?}\comment[id=J]{Does that mean we should not cite the working
papers here?} of agent-based simulation models with contact networks. Examples are
\citet{Basurto2020}, \citet{DelliGatti2020} and \citet{Mellacher2020}.

@@

2. We have an extremely fine grained \replaced[id=K]{representation}{} of schools and
   preschools. We can thus easily model all \replaced[id=K]{schooling policies}{NPIs}
   that have been implemented in Germany in the past nine months. This includes complete
   school closures, phases where only those students whose parents could not find any
   private childcare arrangement could attend, split class approaches for some or all age
   groups and combinations thereof. Moreover, we can account for additional hygiene
   measures whose effect is estimated inside the model.

@@

4. We have an extremely detailed model of PCR and rapid tests with a share of detected
   cases that varies over time and across age groups. We fit estimates of the share of
   known cases from other studies for time periods where those are
   available\comment[id=K]{Citation here?} but also estimate a plausible share of known
   cases after vaccinations become widespread and methodologies that rely on the case
   fatality ratio to identify the share of known cases are not applicable anymore.

 \comment[id=K]{Four is of course plenty but what about weekly recurring contacts or using the google mobility data? -> maybe add these to point 2?}

>>>>

These shortcomings have motivated variations of agent-based simulation models where
moving particles have been replaced by contact networks for households, work and random
contacts. The OpenABM-Covid-19 model by \citet{Hinch2020} and the model by
\citet{Aleta2020} are the closest in spirit to ours.

The OpenABM-Covid-19 model by \citet{Hinch2020} also uses detailed contact networks for
workplaces, schools and households and can evaluate the effect of several NPIs. The main
focus of their application are contact tracing policies. Recently they have also added
support for multiple virus strains and vaccinations.

@@

Moreover, there are several working papers that develop agent-based simulation models
with contact networks in conjunction with economic models. Examples are
\citet{Basurto2020}, \citet{DelliGatti2020} and \citet{Mellacher2020}.

@@

2. We have an extremely fine grained representation of schools and preschools. We can
   thus easily model all schooling policies that have been implemented in Germany in the
   past nine months. This includes complete school closures, phases where only those
   students whose parents could not find any private childcare arrangement could attend,
   split class approaches for some or all age groups and combinations thereof. Moreover,
   we can account for additional hygiene measures whose effect is estimated inside the
   model.

@@

4. We have an extremely detailed model of PCR and rapid tests with a share of detected
   cases that varies over time and across age groups.

5. Our model is designed to combine information from many different data sources.
   Examples are surveys on rapid test demand \citep{Betsch2021}, reaction to test results
   \citep{Betsch2021}, contact diaries \citep{Mossong2008}, share of detected cases
   \citep{Dunkelzifferradar2020} and many more.

### 2021-07-15 src/paper/model/estimated_parameters.tex
Finish most things in section B.

Moreover, we estimate nine different multipliers that reflect how strongly other contacts
are reduced over time. The dates at which we switch between the multipliers usually
coincide with policy changes and is not determined from the \replaced[id=K]{case
numbers}{data}. The only exception to this are slight adjustments to parameters to
incorporate lockdown fatigue (towards the end of a lockdown period) or precautionary
contact reductions (in times of high incidences right before a lockdown is enacted). The
estimated other multipliers are also depicted in Figure~\ref{fig:other_multiplier}.

@@

While we estimate 9 different values for the other contact multiplier, they are not
estimated completely freely. In particular we ensure that the ordering of the parameter
values is consistent with the stringency of policies. For example, the strongest
contact reduction was estimated for January 2021 during where very strict measures
and curfews were in place, whereas the weakest contact reduction was in October 2020
where policies were very lenient.

>>>>

Moreover, we estimate nine different multipliers that reflect how strongly other contacts
are reduced over time. The dates at which we switch between the multipliers usually
coincide with policy changes and is not determined from the case numbers. The only
exception to this are slight adjustments to parameters to incorporate lockdown fatigue
(towards the end of a lockdown period) or precautionary contact reductions (in times of
high incidences right before a lockdown is enacted). The estimated other multipliers are
also depicted in Figure~\ref{fig:other_multiplier}.

@@

While we estimate nine different values for the other contact multiplier, they are not
estimated completely freely. In particular we ensure that the ordering of the parameter
values is consistent with the stringency of policies. For example, the strongest
contact reduction was estimated for January 2021 during where very strict measures
and curfews were in place, whereas the weakest contact reduction was in October 2020
where policies were very lenient.

### 2021-07-15 src/paper/model/initial_conditions.tex

Consider a situation where you want to start a simulation with the beginning set amidst
the pandemic. It means that several thousands of individuals should already have
recovered from the disease, be infectious, symptomatic or in intensive care at the start
of your simulation. Additionally, the sample of infectious people who will determine the
course of the pandemic in the following periods is likely not representative of the
whole population because of differences in behavior (number of contacts,
assortativity), past policies (school closures), etc.. The distribution of courses of
diseases in the population at the begin of the simulation is called initial conditions.

To come up with realistic initial conditions, we match reported infections from official
data to simulated individuals by available characteristics like age and geographic
information. The matching must be done for each day of a longer time frame like a month
to have individuals with possible health states. Then, health statuses evolve until the
begin of the simulation period without simulating infections by contacts. We also
correct reported infections for a reporting lag and scale them up to arrive at the true
number of infections.

>>>>

Consider a situation where you want to start a simulation with the beginning set amidst
the pandemic. It means that several thousands of individuals should already have
recovered from the disease, be infectious, symptomatic or in intensive care at the start
of your simulation. Additionally, the sample of infectious people who will determine the
course of the pandemic in the following periods is likely not representative of the whole
population because of differences in behavior (number of contacts, assortativity), past
policies (school closures), etc.. The distribution of health states in the population at
the beginning of the simulation is called initial conditions.

To come up with realistic initial conditions, we match reported infections from official
data to simulated individuals by age group and county. We use one month of data to
generate initial conditions with in all possible health states. Meanwhile health states
evolve until the beginning of the simulation period without simulating infections by
contacts. We also correct reported infections for a reporting lag and scale them up by
the share of detected cases to arrive at the true number of infections.

### 2021-07-15 src/paper/model/testing.tex

To distribute the correct number of PCR tests among symptomatic and asymptomatic
infections without knowing explicit test demand probabilities, we use the following
approach: First, we calculate the total number of positive PCR tests by multiplying the
number of newly infected individuals with an estimate of the share of detected cases
from the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how
many of these tests should go to symptomatic and asymptomatic individuals from data by
the RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests
are allocated from the pools of symptomatic and asymptomatic infected but not yet tested
individuals.

@@

In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
falsely negative. While the specificity of rapid tests is constant, their sensitivity
strongly depends on the timing of the rapid test relative to the start of infectiousness.
Before the onset of infectiousness the sensitivity is very low (35\%). On the first day
of infectiousness it is much higher (88\%) but still lower than during the remaining
infectious period (92\%). After infectiousness stops, the sensitivity drops to 50 \%.

>>>>

To distribute the correct number of PCR tests among symptomatic and asymptomatic
infections without knowing explicit test demand probabilities, we use the following
approach: First, we calculate the total number of positive PCR tests by multiplying the
number of newly infected individuals with an estimate of the share of detected cases
from the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how
many of these tests should go to symptomatic and asymptomatic individuals from data by
the RKI \citep{ARS2020}. Then, we sample the individuals to which those tests
are allocated from the pools of symptomatic and asymptomatic infected but not yet tested
individuals.

@@

In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
falsely negative. While the specificity of rapid tests is calibrated at 99.4\%
\citep{Bruemmer2021}, their sensitivity strongly depends on the timing of the rapid test
relative to the start of infectiousness \citep{Smith2021}. Before the onset of
infectiousness the sensitivity is very low (35\%). On the first day of infectiousness it
is much higher (88\%) but still lower than during the remaining infectious period (92\%).
After infectiousness stops, the sensitivity drops to 50 \%.
